---
title: "Day 2 Computational Session"
format: html
editor: visual
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse)
```

## Vectors

```{r}
# Create a vector
dogs <- c("teddy", "khora", "waffle", "banjo", "tallie")

# Check the class
class(dogs)
```

```{r}
cats <- c("creamsicle wolf", 14.2, 3L)
class(cats)
# Categorized as character because vector can only be 1 type of class and it will take the most flexible class
```

```{r}
years <- c(20, 40, 16, 2)
class(years)

# typeof function gives you higher resolution of class
typeof(years)

# to get it to return the years vector as an integer class add an L next to each number
```

```{r}
# Want to do a conditional check
# is. <<< asks questions "is this THIS"
# as. <<< gives a command "make this THIS"

is.numeric(years)
is.character(years)
```

```{r}
# Assign names to values in a vector
dog_food <- c(teddy = "purina", khora = "alpo", waffle = "fancy feast")

dog_food

dog_food[2]

dog_food["khora"]
```

```{r}
cars <- c("blue", "orange", "white")
cars[1]

i <- seq(1:2)
cars[i]
```

```{r}
# Assign new value based on position. Called mutability
cars[3] <- "yellow"
cars
```

## Matrices

```{r}
# byrow = false will make it so that it populates values by column instead of by rown
fish_size <- matrix(c(12, 14, 18, 20),ncol = 2, nrow = 2, byrow = FALSE)
fish_size

class(fish_size)
typeof(fish_size)

# Example of indexing: 
fish_size[1,1]
fish_size[2,1]

# What happens if we create a matrix with different types of data?
fish_type <- matrix(c(12, "trout", 18, 13L),ncol = 2, nrow = 2, byrow = FALSE)

class(fish_type)

typeof(fish_type)

fish_type

# leaving column blank (i.e. only having the row indicator) returns all columns for second row
fish_size[2,]

# leaving row blank (i.e. only having the column indicator) returns all rows for first column
fish_size[,1]
```

## Lists

Reminder: A list combines vectors that can be of different length and type

```{r}
urchins <- list("blue", c(1, 2, 3), c("a cat", "a dog"), 5L)
urchins

# using single bracket keeps the class of list, and the double bracket makes the class character
ex_1 <- urchins[3]
ex_2 <- urchins[[3]]
class(ex_1)
class(ex_2)
ex_1[1]
ex_2[1]
```

## Make a named list

```{r}
tacos <- list(topping = c("onion", "cilantro", "guacamole"),
              filling = c("beans", "meat", "veggies"),
              price = c(6.75, 8.25, 9.50))

tacos

# Index list elements by name
tacos$filling

# readr package parsenumber -> only pull out number piece

# return second value of the fillings vector
tacos$filling[2]

```

```{r}
a <- c(1,2,3)
b <- c(3,4,5)

# Run t-test
my_ttest <- t.test(a,b)
my_ttest
class(my_ttest)
my_ttest$p.value
```

The price of a veggie burrito with guacamole is 'r tacos\$price\[2\]'

In r markdown add information in text in a way that's reproducible.

'r my_list\$dogs\[2\]'

## Data frames

3 ways to create a data frame

```{r}
# Create 3 separate vectors and then combine them into a data frame using data.frame function
# vectors in data frome need to have the same length
topping <- c("onion", "cilantro", "guacamole")
filling <- c("beans", "meat", "veggies")
price <- c(6.75, 8.25, 9.50)

df_ex_1 <- data.frame(topping, filling, price)
class(df_ex_1)

# combine all commands into single function
df_ex_2 <- data.frame(topping = c("sour cream", "salsa", "cheese"),
                      filling = c("chicken", "cheese", "cabbage"),
                      price = c(12.50, 13.75, 15.00))

# make data frome/tibble, use the tidyverse package, a way to write it out how it will look in the data frame
df_ex_3 <- tribble(
  ~topping, ~filling, ~price,
  "avocado", "steak", 13.25,
  "radish", "fish", 10.40,
  "salsa", "veggies", 9.75
)

class(df_ex_3)

df_ex_3

df_ex_3$price[3]
```

Because a data frame is a list allows us to have different classes for each column
